### 2020-09-03 LaTeX/sections/paper.tex
Edits

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =
\section{Cryptocurrency Exchanges}

@@

% = = = = = = = = = = = = = = = = =
\subsection{Centralized Exchanges (CEX)}\label{sec:CEX}

@@

% = = = = = = = = = = = = = = = = =
\subsection{Decentralized Exchanges (DEX)}

@@

\section{Order Books}

@@

\section{Call Market}\label{sec:callmarket}

\subsection{Priority Queues}

A priority queue can be implemented using many of the data structures that differ in performance and costs. In order to choose the appropriate data structure, we implemented and examined five different generic priority queues that support four main functions (see Table~\ref{tab:PQ_API}). We discuss each implementation below.

% = = = = = = = = = = = =Operations for a generic Priority Queue Table = = =  = = = = = = = = = = = %
\begin{table}[]
\centering
\begin{tabular}{|c|c|}
\hline

@@

@@

\item\textbf{{Heap with Dynamic Array.}} A heap is a type of binary tree data structure that comes in two forms of a (i) Max-Heap and (ii) Min-Heap. All the nodes of tree are in a specific order and the root of the tree always represents the highest priority item of the data structure (the largest and smallest values in the Max-Heap and Min-Heap respectively). We implemented a priority queue with a heap that itself is represented by a Solidity dynamically-sized array.

@@

\item\textbf{{Mapping with Keys Stored in a Heap.}} This priority queue stores items inside a Solidity mapping that maps them to mapping keys (256-bit integers). These keys are then stored in a sorted heap (based on the value they are mapped to) that is implemented with a dynamically-sized array.

@@

\item\textbf{{Linked List.}} Here we used a sorted singly linked list to implement the priority queue. A linked list is a linear data structure made of a chain of items (also known as nodes) where each item contains a value and a pointer to the next node in the list. In an object-oriented programming language a linked list can be easily implemented with a recursive struct; struct that contains a member variable in its own type. However, Solidity does not support recursive structs because the size of the struct has to be finite. So in our implementation, each node is represented by a smart contract that contains the data and a pointer of its own type to its next smart contract in the list.

@@

\item\textbf{{Linked List with Mapping.}} Here we used Solidity mapping to implement a sorted doubly linked list priority queue where every node points to it previous \textbf{and} next node in the list. The Linked List entries are first wrapped in an struct data type that contains (i) the unique id of that element, (ii) the unique id of its previous item in the list, and (iii) the unique id of its next item in the list. These structs are then stored in the mapping indexed by item id.

@@

In the next step, we used Javascript for testing by leveraging the Mocha testing framework and performed performance tests on the five priority queues. We entered 50 pseudo-random integers between 1 and 200 to the priority queues and obtained the associated gas costs for 200 transactions. Figure~\ref{fig:random_insertion} represents the gas costs variations for entering 50 unsigned integers to the five priority queues; the x-axis is the place in line (\eg the10th number entered in the priority queue) and the and y-axis is the cost of that transaction in gas.

@@

\begin{figure}[htb!p]
\centering
\includegraphics[width=1\textwidth]{fig/random_insertion.png}
\caption{\footnotesize{Gas costs for entering 50 items into five different priority queues.}  \label{fig:random_insertion}}
\end{figure}

@@

Then, we called the \texttt{Dequeue()} function which iteratively removes the highest priority number from the priority queues (until the data structure is empty). The computational costs for dequeuing 50 unsigned integers in each priority queue are outlined in Table~\ref{tab:PQUnitTests} where (\full) indicates the full refund and (\prt) means the amount of refund is bounded. These tests are performed using the current Ethereum gas metrics (block gas limit $=$11,741,495 and 1 gas $=$ 56 gwei)~\footnote{https://ethstats.net/}. The second column of the table shows the net gas consumption (the gasUsed value derived from transaction receipts) for removing 50 integers from each priority queues.

%Truffle gas limit for the tests: 11741495

\begin{table}[]
\centering
\begin{tabular}{|c|c|c|}
\hline

\textbf{\thead{Priority Queue}}    & \textbf{\thead{Net Cost in Gas\\ (\texttt{gasUsed})}}             & \textbf{\thead{Gas Refunds from Deleting a Storage\\(Manually Calculated)}} \\ \hline

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Dynamic Array}}         				& 2,518,131            		      & 750,000                       \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Static Array}}           				& 1,385,307                             & 750,000                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Mapping with \\ keys stored \\ in Heap}} 		& 2,781,684                            & 1,500,000                     \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{Linked List}                       							& 557,085               	           & 1,200,000                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Linked List with \\ Mapping}}          				& 731,514              	     	  &  3,765,000                       \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %

% Below is the same table but with "Full, Partial,None" for refunds

\begin{table}[]
\centering
\begin{tabular}{|c|c|c|}
\hline

\textbf{\thead{Priority Queue}}    & \textbf{\thead{Net Cost in Gas\\ (\texttt{gasUsed})}}             & \textbf{\thead{Gas Refunds from Deleting a Storage}} \\ \hline

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Dynamic Array}}         				& 2,518,131            		      & \full                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Heap with \\ Static Array}}           				& 1,385,307                             & \full                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Mapping with \\ keys stored \\ in Heap}} 		& 2,781,684                            & \full                     \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{Linked List}                       							& 557,085               	           & \full                      \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %
	\textbf{\thead{Linked List with \\ Mapping}}          				& 731,514              	     	  &  \full                       \\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = %

Note that in order to urge miners to process smart contracts with refunds, the accumulated gas refund can never exceed half the gas used up during computation~\cite{wood2014ethereum}. So at the end of a successful transaction, the amount of gas in the refund counter (capped at half the net gas used) is returned to the caller. Let's take a closer look at the last row of Table~\ref{tab:PQUnitTests}; gas that has been used for dequeuing 50 integers from the linked list with mapping priority queue is 731,514 and the gas limit has been set up to 11,741,495. Firstly, the caller will receive $(11,741,495 - 731,514) = 11,009,981$ units of unused gas. Also, the amount of refund we manually calculated based on the storage deletions is 3,765,000, and since this number is greater than half of he amount of gas that has been used in the contract ($3,765,000 > 731,514/2$), the refund generated will be $731,514/2 = 365,757$. So the caller of this transaction will eventually receive is $11,009,981 + 365,757 = 11,375,738$ units of gas.

%& \textbf{\thead{Net Cost\\in Gas\\ (\texttt{gasUsed})}}  
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

@@

Based on our tests results, the \textit{linked list} and \textit{linked list with mapping} priority queues show a better performance in compared to the different heap implementations. In the next section, we explore the differences between the two different implementations of the linked list priority queue.

\subsection{Linked List \textit{vs.} Linked List with Mapping}\label{sec:LLvsLLMapping}

There are two main ways to implement each of the \textit{linked list} and \textit{linked list with mapping} priority queues. In this section, we perform a comprehensive comparison between these two with the purpose of choosing the proper data structure for implementing the call market.

@@

@@

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %
\begin{table}[]
\begin{tabular}{|c|c|c|c|c|}
\hline
\multirow{2}{*}{\textbf{Gas Metrics}}                        							& \multicolumn{2}{c|}{\textbf{Linked List}}           						& \multicolumn{2}{c|}{\textbf{Linked List with Mapping}}     					\\ \cline{2-5} 
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %
                                                                								& with \texttt{SELFDESTRUCT} 		& without \texttt{SELFDESTRUCT}                	& with \texttt{DELETE} 		& without \texttt{DELETE}              		\\ \hline                                                                                                                                                                                                                                                          % = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %
                  \textbf{\thead{Net Cost in Gas\\(\texttt{gasUsed})}}                        				& 557,085                    			& 721,370                     			& 731,514                     		& 334,689                      				\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %
 \textbf{\thead{Gas Refund \\from Deleting a Storage\\(Manually Calculated)}}            			& 1,200,000                 					& 0                					& 3,765,000                    		& 765,000                       				\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %

% Below is the same table but with "Full, Partial,None" for refunds

\begin{table}[]
\begin{tabular}{|c|c|c|c|c|}
\hline
\multirow{2}{*}{\textbf{Gas Metrics}}                        							& \multicolumn{2}{c|}{\textbf{Linked List}}           						& \multicolumn{2}{c|}{\textbf{Linked List with Mapping}}     					\\ \cline{2-5} 
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %
                                                                								& with \texttt{SELFDESTRUCT} 		& without \texttt{SELFDESTRUCT}                	& with \texttt{DELETE} 		& without \texttt{DELETE}              		\\ \hline                                                                                                                                                                                                                                                          % = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %
                  \textbf{\thead{Net Cost in Gas\\(\texttt{gasUsed})}}                    & 557,085                    			& 721,370                     			& 731,514                     		& 334,689                      				\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %
 \textbf{\thead{Gas Refund \\from Deleting a Storage}}            			& \full                 					& $\times$                 					& \full                    		& \prt                       				\\ \hline
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = = = = = = = = = = =  = = = = = =  = = = = = = %

Based on Table~\ref{tab:LLvsLLMapping}, the linked list with \texttt{SELFDESTRUCT} appears to be a good candidate for the call market implementation. However, as it can be seen in Figure~\ref{fig:random_insertion}, the associated cost for inserting elements into this priority queue is significantly greater than the linked list with mapping as insertion is identical to creating new contracts. Accordingly, we choose to implement the call market with the linked list with mapping that uses the  the \texttt{DELETE} method to clear up the chain every time the highest priority element is removed from the list. In overall this priority queue introduces a moderate gas cost for both insertion (\ie order submission) and removal (\ie matching the orders).

@@

\section{Experimentations}

>>>>

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =
\section{Preliminaries}

@@

\subsection{Blockchain}

* transparency, front-running
* slow but is becoming more efficient over time, expensive
* Gas model: gas limit, how much, per transaction vs per block, Solidity/EVM quirks: gas refunds

Note that in order to urge miners to process smart contracts with refunds, the accumulated gas refund can never exceed half the gas used up during computation~\cite{wood2014ethereum}. So at the end of a successful transaction, the amount of gas in the refund counter (capped at half the net gas used) is returned to the caller.

\subsection{Exchanges}

% = = = = = = = = = = = = = = = = =
\subsubsection{Centralized Exchanges (CEX)}\label{sec:CEX}

@@

% = = = = = = = = = = = = = = = = =
\subsubsection{Decentralized Exchanges (DEX)}

@@

\subsection{Order Books}

@@

\section{Priority Queue}\label{sec:pq}

% = = = = = = = = = = = =Operations for a generic Priority Queue Table = = =  = = = = = = = = = = = %
\begin{table}[t]
\centering
\begin{tabular}{|c|c|}
\hline

@@

In designing \cm, performance within Ethereum's limited gas model became the main bottleneck. Within a call market, recording trades and accessing them to close the market is the most time consuming step. For this reason, we decided to do a deep study of the best data-structure for this specific task. While data structures are well studied for many languages, Solidity/EVM have their own quirks (\eg gas refunds, a relatively cheap mapping data structure, and no actual support for object oriented programming) that make them difficult to assess which will perform best without actually deploying and evaluating each variant.

When closing a call market, the orders are examined in order: highest to lowest price for bids, and lowest to highest price for asks. It most circumstances, the market closing algorithm does not have to consider any deeper bids/asks from the list when choosing whether the current best bid and ask can be fulfilled. The only exception is in the case of a tie on price, which we return to later. For this reason, the ideal data structure for storing bids/asks is a \textit{priority queue} (see Table~\ref{tab:PQ_API}) where each order's priority is its price. Specifically, we use two PQs---one for bids where the highest price is the highest priority, and one for asks where the lowest price is the highest priority.

There are numerous ways of implementing a PQ. A PQ has an underlying list---common options include a static array, dynamic array, and linked list. The most expensive operation is keeping the data sorted---common options include (i) sorting during each enqueue, (ii) sorting for each dequeue, or (iii) splitting the difference by using a heap as the underlying data structure. Respectively, the time complexities are (i) linear enqueue and constant dequeue, (ii) constant enqueue and linear dequeue, and (iii) logarithmic enqueue and logarithmic dequeue. This said, Solidity/EVM has its own specific considerations to take into account. As closing the market is very expensive with any PQ, we rule out using (ii) as fully sorting while dequeuing would be prohibitive. We experiment with the following 5 options for (i) and (iii):

@@

\item\textbf{{Heap with Dynamic Array.}} A heap is a type of binary tree data structure that comes in two forms of a (i) Max-Heap and (ii) Min-Heap. All the nodes of tree are in a specific order and the root of the tree always represents the highest priority item of the data structure (the largest and smallest values in the Max-Heap and Min-Heap respectively). We implemented a priority queue with a heap that stores its data in a dynamically-sized array.

@@

\item\textbf{{Heap with Mapping.}} In the above implementations, the heap stores the entire order (as a struct) in the heap. In this variant, we store the order struct in a Solidity mapping and store the mapping's key in the heap.

@@

\item\textbf{{Linked List.}} In this variant, we insert a new element into its correct position (based on its price) when running enqueue. The PQ itself stores elements in a linked list (enabling us to efficiently insert a new element between two existing elements). Solidity is described as object-oriented but the equivalent of an object is an entire smart contract. Therefore an object-oriented linked list must either (i) create each node in the list as a struct---but this is not possible as Solidity does not support recursive structs---or (ii) make every node in the list its own contract. The latter option seems wasteful and unusual, but we try it out anyways. Thus each node is its own contract and contains the order data and a pointer to the address of the next contract in the list.

@@

\item\textbf{{Linked List with Mapping.}} Finally, we try a variant of a linked list using a Solidity mapping. The value of the mapping is a struct with order data and the key of the next (and previous) node in the list. The contract stores the key of the first node (head) in the list.

@@

\subsection{Priority Queue Evaluation}

\subsubsection{Enqueue performance.} We implemented, deployed, and tested each priority queue using Truffle and Ganache.\footnote{Github: Link removed for anonymity.} We tried a variety of tests (including testing the full call market with each varient) with consistent results in performance. A simple test to showcase the performance profile is shown in Figure~\ref{fig:random_insertion}. We simply enqueue 50 integers chosen at random from a fixed interval in each PQ variant. The bigger the PQ gets, the longer enqueue takes---a linear increase for the linked list variants, and logarithmic for the heap variants.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.8\textwidth]{fig/random_insertion.png}
\caption{\footnotesize{Gas costs for enqueuing random items into five different priority queue variants. For the x-axis, a value of 10 indicates it is the 10th number entered in the priority queue. The y-axis is the cost of enqueue in gas.}  \label{fig:random_insertion}}
\end{figure}

\input{tables/dequeue.tex}

@@

\subsubsection{Dequeue performance.} For each PQ variant storing 50 random integers, the \texttt{Dequeue()} function is iterated until the data structure is empty. The total gas cost for fully dequeuing the priority queue variants is outlined in Table~\ref{tab:PQUnitTests}. These tests are performed using the following Ethereum gas metrics: block gas limit $=$11,741,495 and 1 gas $=$ 56 gwei.\footnote{\[July 2020:\] https://ethstats.net/} Dequeuing removes data from the contract's storage. Recall this results in a gas refund. Based on our manual estimates (EVM does not expose the refund counter), every variant receives the maximum gas refund possible (\ie half the total cost of the transaction). In other words, each of them actually consumes twice the \texttt{gasUsed} amount in gas before the refund, however none of them are better or worst based on how much of a refund they generate.

\subsubsection{Decision.} Based on enqueing, the heap variants are the cheapest in terms of gas, while based on dequeuing, the link list variants are cheapest. This is in accordance with the theoretical worst-case time complexity for each. However, (i) the linked list variants are materially cheaper than the heap variants at dequeuing, and (ii) dequeuing in a call market must be done as a batch, whereas enqueuing is paid for one at a time by the trader submitting the order, and (iii) Ethereum will not permit more than hundreds of orders so the asymptotic behaviour is not a significant factor. For these reasons, we suggest using a linked list variant for this specific application. As it can be seen in Figure~\ref{fig:random_insertion}, the associated cost for inserting elements into a linked list PQ is significantly greater than the linked list with mapping, as each insertion cause the creation of a new contracts. Accordingly, we choose to implement the call market with the linked list with mapping. Overall this PQ balances a moderate gas cost for insertion (\ie order submission) with one for removal (\ie matching the orders).

@@

\subsection{Cost/Benefit of cleaning up after yourself}
\label{sec:LLvsLLMapping}
\label{sec:gasrefund}

\input{tables/cleaning.tex}

@@

% With 50 selfdestructs

@@

 that uses the  the \texttt{DELETE} method to clear up the chain every time the highest priority element is removed from the list. In

@@

\section{Call Market Design}
